### 2018-04-26 paper/source.tex
Created TODO list and some other support files.
Namely: .gitignore, source.tex and run.sh

%% LyX 2.1.4 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage[T1]{fontenc}
\usepackage{listings}
\usepackage[usenames,dvipsnames]{xcolor}

%%
%% Julia definition (c) 2014 Jubobs
%%
\lstdefinelanguage{Julia}%
  {morekeywords={abstract,break,case,catch,const,continue,do,else,elseif,%
      end,export,false,for,function,immutable,import,importall,if,in,%
      macro,module,otherwise,quote,return,switch,true,try,type,typealias,%
      using,while},%
   sensitive=true,%
   alsoother={$},%
   morecomment=[l]\#,%
   morecomment=[n]{\#=}{=\#},%
   morestring=[s]{"}{"},%
   morestring=[m]{'}{'},%
}[keywords,comments,strings]%

\lstset{%
    language=julia,
    breaklines=true,
    frame=lines,
    basicstyle={\tiny\ttfamily},
    numbers=left,
    stepnumber=10,
    numberstyle={\tiny},
    tabsize=4,
    commentstyle={\color{gray}\itshape},
    keywordstyle={\color{blue}},
    stringstyle={\color{teal}},
    showstringspaces=false
}

\makeatother

\usepackage{babel}
\usepackage{listings}
\begin{document}

\title{Island Speciation Model Source Code}

\author{Ludwig Leidinger, Daniel Vedder}

\maketitle
\lstinputlisting[caption={islandsim.jl}]{../islandsim.jl}

\lstinputlisting[caption={extractinfo.jl}]{../extractinfo.jl}

\lstinputlisting[caption={getdata.jl}]{../getdata.jl}

\lstinputlisting[caption={GeMM.jl}]{../src/GeMM.jl}

\lstinputlisting[caption={types.jl}]{../src/types.jl}

\lstinputlisting[caption={init.jl}]{../src/init.jl}

\lstinputlisting[caption={mainprocs.jl}]{../src/mainprocs.jl}

\lstinputlisting[caption={subprocs.jl}]{../src/subprocs.jl}

\lstinputlisting[caption={io.jl}]{../src/io.jl}
\end{document}